### 2024-03-22 doc/sections/mav-extensions.tex
Un-agda files

\section{Extensions}\label{sec:mav-extensions}

\subsection{Additive Units}

\subsection{Multiplicative-Additive Linear Logic}

\subsection{Display Calculus}

Cut-elimination for a Display calculus?