### 2024-03-26 doc/sections/mav-semantics.tex
[ tex ] done most of the semantics section

\section{Semantic Models for MAV}\label{sec:mav-semantics}

\newcommand{\LowerSet}[1]{\widehat{#1}}
\newcommand{\Day}[1]{\mathop{\widehat{#1}}}
\newcommand{\ClosedLowerSet}[1]{\widehat{#1}^+}
\newcommand{\ClosedDay}[1]{\mathop{\widehat{#1}^+}}
\newcommand{\Chu}{\mathrm{Chu}}
\newcommand{\op}{\mathrm{op}}

\bob{Short introduction to the section here}

\begin{itemize}
\item We construct algebraic models of MAV
\item MAV-algebras are algebras that MAV is sound for. In some
  sense, this is straightforward, but we identify precisely the
  structure that is needed.
\item The key structure is duoidality, or entropy.
\item We develop some model theory here, and show that MAV-algebras
  can be induced by the weaker structure of MAV-frames. This is
  essential for the cut-free elimination proof.
\end{itemize}

\subsection{Pomonoidal, $*$-autonomous, and Duoidal Structure on Partial Orders}

\begin{definition}
  A \emph{partial order monoid (pomonoid)} $(\bullet, i)$ on a poset
  $(A, \leq)$ comprises a binary operator $\bullet : A \times A \to A$
  that is monotone in both arguments and an element $i \in A$ such
  that the usual monoid laws hold. A \emph{commutative pomonoid} is a
  pomonoid where additionally $x \bullet y = y \bullet x$.
\end{definition}

\begin{definition}
  \bob{residuation}
\end{definition}

The definition of $*$-autonomous category is due to Barr
\cite{Barr_1979}. For our purposes, we need the partial order
analogue, also called a CL algebra by Troelstra
\cite{Troelstra92:lll}.

\begin{definition}
  A \emph{$*$-autonomous partial order} is a structure
  $(A, \leq, \otimes, I, \lnot)$ where $(\otimes, I)$ is a pomonoid on
  $(A, \leq)$ and $\lnot : A^\op \to A$ is an anti-monotone and
  involutive operator on $A$, together satisfying
  $x \otimes y \leq \lnot z$ iff $x \leq \lnot (y \otimes z)$.  A
  $*$-autonomous partial order satisfies \emph{mix} if $\lnot I = I$.
\end{definition}

\begin{remark}
  The structure of a $*$-autonomous partial order has a number of
  immediate consequences, but we leave description of these until
  after the definition of MAV-algebra in Definition
  \ref{defn:mav-algebra}.
\end{remark}

FIXME: cite that combinatorics book.

\bob{need to define \emph{entropy} and \emph{entropic} somewhere. I
  think we can have that one pomonoid is \emph{entropic} over another,
  and the structure of having two monoids where one is entropic over
  the other is called duoidal structure.}

\begin{definition}
  Two pomonoids $(\bullet, i)$ and $(\lhd, j)$ on a partial order
  $(A, \leq)$ are in a \emph{duoidal} relationship if the following
  inequalities hold:
  \begin{enumerate}
    \item $(w \lhd x) \bullet (y \lhd z) \leq (w \bullet y) \lhd (x \bullet z)$
    \item $j \bullet j \leq j$
    \item $i \leq i \lhd i$
    \item $i \leq j$
  \end{enumerate}
\end{definition}

\begin{remark}
  In the case when the two pomonoids share a common unit the last
  three conditions for a duoidal relationship are automatically
  satisfied.
\end{remark}

\begin{remark}
  If $\bullet$ is a join, or $\lhd$ is a meet, then all the conditions
  for a duoidal relationship are automatically met for any other
  monoid. \bob{fwd ref to where we use this.}
\end{remark}

\begin{remark}
  The duoidal relationship is not symmetric.
\end{remark}

\subsection{MAV-algebras}

\bob{Change $;$ to $\lhd$}

\begin{definition}\label{defn:mav-algebra}
  An \emph{MAV-algebra} is a structure
  $(A, \leq, \otimes, ;, I, \lnot)$ with the following properties:
  \begin{enumerate}
  \item $(A, \leq, \otimes, I, \lnot)$ is $*$-autonomous and satisfies \emph{mix}.
  \item $(A, \leq, ;, I)$ is a pomonoid.
  \item $;$ is self dual: $\lnot (x ; y) = (\lnot x); (\lnot y)$.
  \item $(\otimes, I)$ is duoidal over $(;, I)$.
  \item $(A, \leq)$ has binary meets, which we write as $x \with y$.
  \end{enumerate}
\end{definition}

\begin{proposition}\label{prop:mav-algebra-consequences}
  Let $(A, \leq, \otimes, ;, I, \lnot)$ be a MAV-algebra.
  \begin{enumerate}
    \item There is another pomonoid structure $(\parr, I)$ on
          $(A, \leq)$, defined as
          $x \parr y = \lnot(\lnot x \otimes \lnot y)$.
    \item $(\otimes, I)$ and $(\parr, I)$ are linearly distributive
          \cite{cockett}:
          $x \otimes (y \parr z) \leq (x \otimes y) \parr z$.
    \item $(A, \leq)$ has binary joins, given by
          $x \oplus y = \lnot (\lnot x \with \lnot y)$.
    \item $\oplus$ distributes over $\otimes$:
          $x \otimes (y \oplus z) = ((x \otimes y) \oplus (x \otimes z)$.
    \item $\with$ distributes over $\parr$:
          $(x \parr z) \with (y \parr z) = (x \with y) \parr z$.
    \item $;$ is entropic over $\parr$:
          $(w \parr x); (y \parr z) \leq (w;y) \parr (x;z)$.
    \item $;$ is entropic over $\with$:
          $(w \with x); (y \with z) \leq (w ; y) \with (x; z)$.
    \item $\oplus$ is entropic over $;$:
          $(w ; x) \oplus (y ; z) \leq (x \oplus y) ; (x \oplus z)$.
  \end{enumerate}
\end{proposition}

\newcommand{\sem}[1]{\llbracket #1 \rrbracket}

\begin{definition}\label{defn:mav-interpretation}
  Let $\mathit{At}$ be a set of atomic propositions. Given an
  MAV-algebra $(A, \leq, \otimes, \lhd, I, \lnot)$ and valuation
  $V : \mathit{At} \to A$, define the interpretation of MAV Formulas
  as follows: $\sem{\vUnit} = I$, $\sem{a} = V(a)$,
  $\sem{\vDual a} = \lnot V(a)$,
  $\sem{\vP\vTens\vQ} = \sem{\vP} \otimes \sem{\vQ}$,
  $\sem{\vP\vParr\vQ} = \sem{\vP} \parr \sem{\vQ}$,
  $\sem{\vP\vSeq\vQ} = \sem{\vP};\sem{\vQ}$,
  $\sem{\vP\vWith\vQ} = \sem{\vP} \with \sem{\vQ}$, and
  $\sem{\vP\vPlus\vQ} = \sem{\vP} \oplus \sem{\vQ}$.
\end{definition}

\begin{lemma}
  For all $\vP$, $\sem{\vDual{\vP}} = \lnot \sem{\vP}$.
\end{lemma}

\begin{theorem}\label{thm:soundness}
  The interpretation defined in Definition
  \ref{defn:mav-interpretation} is sound: for all formulas $P$, if
  $P \longrightarrow^* I$, then $I \leq \sem{P}$.
\end{theorem}

\begin{proof}
  Each of the required inequalities has been established in
  Proposition \ref{prop:mav-algebra-consequences}.
\end{proof}

\subsection{MAV-frames}

To prove completeness of the cut-free fragment of SMAV, we will
construct a particular MAV-algebra from the formulas and cut-free
proofs. Since cut-free proofs do not a priori have all the necessary
structure for an MAV-algebra, we develop a procedure to construct
MAV-algebra from the lighter requirements of an MAV-frame, defined in
this section. MAV-frames have an intuitive reading as a kind of
process algebra, so we use process algebra-like notation. In the
following sections, we show how every MAV-frame induces an
MAV-algebra, and in Section XXX we will show that the MAV-algebra
constructed from the cut-free proof MAV-frame allows us to prove that
all proofs in SMAV have cut-free normal forms.

\begin{definition}
  An \emph{MAV-frame} is a structure
  $(F, \leq, I, \parallel, \lhd, +)$ where $(F, \leq)$ is a partial
  order, $(F, \leq, I, \parallel)$ is a commutative pomonoid,
  $(F, \leq, I, \lhd)$ is a pomonoid, $+$ is a binary monotone
  function on $(F, \leq)$, and these data satisfy the following
  inequalities:
  \begin{enumerate}
    \item $(w \lhd x) \parallel (y \lhd z) \leq (w \parallel y) \lhd (x \parallel z)$
    \item $(x + y) \parallel z \leq (x \parallel z) + (y \parallel z)$
    \item $(w \lhd x) + (y \lhd z) \leq (w + y) \lhd (x + z)$
    \item $I + I \leq I$
  \end{enumerate}
\end{definition}

\begin{remark}
  An MAV-frame is essentially two duoidal relationships and a
  distributivity law.
\end{remark}

\begin{remark}
  MAV-frames have a intuitive reading as a CCS-like process algebra
  (see Milner \cite{milner89} for an introduction to CCS). If we
  assume the existence of a collection of ``action'' elements
  $a \in F$ and their duals $\overline{a} \in F$, satisfying
  $a \parallel \overline{a} \leq I$, then we can read the constructs
  of an MAV-frame as parallel composition, sequential composition, and
  choice. The ordering is interpreted as a reduction relation. An
  interesting avenue for future work would be to discover to what
  extent MAV can be thought of as a logic for processes in this
  process algebra.
\end{remark}

\begin{proposition}
  The \emph{cut-free MAV-frame} $\textsc{CFMav}$ is the partial order
  arising as the quotient of the preorder formed from the formulas of
  MAV and $P \leq Q$ if $P \longrightarrow^* Q$. Define
  $P \parallel Q = P \vParr Q$, $P \lhd Q = P \vSeq Q$,
  $P + Q = P \vWith Q$ and $I = \vUnit$. The required (in)equalities
  follow directly from the definition of $\longrightarrow^*$.
\end{proposition}

\begin{remark}
  The construction of the MAV-frame \textsc{CFMav} does not use the
  $\otimes$ and $\oplus$ structure of MAV directly. We will see in
  Section \ref{sec:mav-cut-elimination} that this structure is
  recovered by duality from the other connectives by the constructions
  in the rest of this section. \bob{Which one is ``positive'' again?}
\end{remark}

\subsection{Constructing MAV-algebras from MAV-frames}

We construct MAV-algebras from MAV-frames in a three step process. In
Section \ref{sec:lower-sets}, we use lower sets and the Day
construction to add meets, joins and residuals for pomonoids to a
partial order. This construction creates joins freely, so we restrict
to $+$-closed lower sets in Section \ref{sec:closed-lower-sets}, which
separates the Day construction of pomonoids into two separate
cases. Finally, we create the $*$-autonomous structure using the Chu
construction in Section \ref{sec:chu}. We maintain the duoidal
structure through each construction.

\subsubsection{Lower Sets and Day pomonoids}
\label{sec:lower-sets}

\begin{definition}
  Given a partial order $(A, \leq)$, the set of lower sets
  $\LowerSet{A}$ consists of subsets $F \subseteq A$ that are
  down-closed: $x \in F$ and $y \leq x$ implies $y \in F$. Lower sets
  are ordered by inclusion. Define the embedding
  $\eta : A \to \LowerSet{A}$ as $\eta(x) = \{ y \mid y \leq x \}$.
\end{definition}

\begin{proposition}
  For any $(A, \leq)$, the function $\eta$ is monotone, and
  $(\LowerSet{A}, \subseteq)$ has meets and joins given by
  intersection and union respectively.
%
%  \bob{and $\eta$ preserves any meets that exist}
\end{proposition}

\begin{proposition}\label{prop:day-construction}
  If $(\bullet, i)$ is a pomonoid on $(A, \leq)$, then there is a
  corresponding \emph{Day pomonoid} $(\Day{\bullet}, \Day{i})$ on
  $\LowerSet{A}$ defined as
  $F \Day\bullet G = \{ z \mid z \leq x \bullet y, x \in F, y \in G
    \}$ and $\Day{i} = \eta(i)$. Moreover:
  \begin{enumerate}
    \item $(\Day{\bullet}, \Day{i})$ is left and right residuated: there
          exists a function
          $\Day{\rightblackspoon} : \LowerSet{A} \times \LowerSet{A} \to
            \LowerSet{A}$ such that ... FIXME
    \item If $(\bullet, i)$ is a commutative pomonoid, then so is
          $(\Day{\bullet}, \Day{i})$. In this case, the residuals coincide.
  \end{enumerate}
  \bob{Could also say that $\eta$ preserves the monoids. Does this
    come up later?}
\end{proposition}

\begin{remark}
  This is the Day construction \cite{day} restricted to the case of
  partial orders.
\end{remark}

\begin{remark}
  When $(A, \leq)$ is an MAV-frame, by \ref{prop:day-construction} we
  obtain two pomonoids $(\Day{\parallel}, \Day{I})$ and
  $(\Day{\lhd}, \Day{I})$ on $\LowerSet{A}$. Moreover, the next
  proposition states that the duoidal relationship between these
  monoids is preserved by the Day construction:
\end{remark}

\begin{proposition}\label{prop:lower-set-duoidal}
  If $(\bullet, i)$ is duoidal over $(\lhd, j)$ then
  $(\Day{\bullet}, \Day{i})$ is duoidal over $(\Day{\lhd}, \Day{j})$.
\end{proposition}

\subsubsection{$+$-closed Lower Sets}
\label{sec:closed-lower-sets}

In the phase semantics, formulas are interpreted as elements that are
fixed points of a closure operator defined by a double negation with
respect to the monoid on the original frame. This closure operator
generates a partial order of ``facts'' whose meets and joins exactly
correspond to the syntactic ones when the original monoid is derived
from the proofs. As we mentioned in the introduction, the presence of
a self-dual operator in BV means that we cannot use double negation
closure, and we have to proceed more deliberately to preserve
join-like structure in an MAV-frame when building MAV-algebras. We do
this by defining $+$-closed lower sets as those that are closed under
finite $+$-combinations of their elements. This leads to a closure
operator on lower sets that allows us to immediately deduce that
$+$-closed lower sets form a lattice. We also preserve the
Day-pomonoids from lower sets, but in two different ways, depending on
how the original pomonoid interacts with $+$. In Proposition
\ref{prop:closed-monoid-distrib} we handle pomonoids that distribute
over $+$, and in Propositional \ref{prop:closed-monoid-duoidal} we
handle pomonoids that are entopic under $+$. We need these
constructions to lift the $\parallel$ and $\lhd$ pomonoids from
MAV-frames to $+$-closed lower sets. Finally in this section, we show
that duoidal structure on lower sets from Proposition
\ref{prop:lower-set-duoidal} is preserved in $+$-closed lower sets.

For this section, we assume that $(A, \leq)$ is a partial order with a
monotone binary operation $+ : A \times A \to A$ (we do not assume
that $+$ is a join or even a pomonoid.)

\begin{definition}
  A lower set $F \in \LowerSet{A}$ is \emph{$+$-closed} if $x \in F$
  and $y \in F$ imply $x + y \in F$. $+$-closed lower sets are ordered
  by subset inclusion and form a partial order
  $(\ClosedLowerSet{A}, \subseteq)$.
\end{definition}

\begin{proposition}
  Let $U : \ClosedLowerSet{A} \to \LowerSet{A}$ be the ``forgetful''
  function that forgets the $+$-closed property. There is a monotone
  function $\alpha : \LowerSet{A} \to \ClosedLowerSet{A}$ such that
  for all $F \in \ClosedLowerSet{A}$, $\alpha(U F) = F$ and for all
  $F \in \LowerSet{A}$, $F \subseteq U (\alpha F)$.
\end{proposition}

\begin{proof}
  To define $\alpha$, we close lower sets under all
  $+$-combinations. To this end, for $F \in \LowerSet{A}$, define
  $\mathrm{ctxt}(F)$, the set of all $+$-combinations of $F$
  inductively built from constructors
  $\mathsf{leaf} : F \to \mathrm{ctxt}(F)$ and
  $\mathsf{node} : \mathrm{ctxt}(F) \times \mathrm{ctxt}(F) \to
  \mathrm{ctxt}(F)$. We define the \emph{sum} of a context as
  $\mathit{sum}(\mathsf{leaf}~x) = x$ and
  $\mathit{sum}(\mathsf{node}(c,d)) = \mathit{sum}(c) +
  \mathit{sum}(d)$. Now define:
  $\alpha(F) = \{ x \mid c \in \mathrm{ctxt}(F), x \leq
  \mathit{sum}(c) \}$. This is $+$-closed, by taking the
  $\mathsf{node}$ combination of contexts. $\alpha \circ U$ is
  idempotent because $\alpha$ does not introduce any elements to lower
  sets that are already closed. For arbitrary lower sets $F$,
  $F \subseteq U(\alpha F)$ by the $\mathsf{leaf}$ constructor.
\end{proof}

\begin{remark}
  By this proposition, $U \circ \alpha$ is a closure operator on
  $\LowerSet{A}$ \cite{davey-priestley}, and the closed elements are
  those of $\ClosedLowerSet{A}$. The next proposition is standard for
  showing that meets and joins exist on the closed elements for some
  closure operator.
\end{remark}

\begin{proposition}
  $(\ClosedLowerSet{A}, \subseteq)$ has all meets and joins. In the
  binary case, meets are defined by intersection and joins are defined
  by $F \lor G = \alpha (U F \cup U G)$.
\end{proposition}

\begin{remark}
  Note that $\alpha$ does not preserve meets.
\end{remark}

\begin{proposition}\label{prop:closed-monoid-distrib}
  For a commutative pomonoid $(\bullet, i)$ on $(A, \leq)$ that
  distributes over $+$
  ($(x + y) \bullet z \leq (x \bullet z) + (y \bullet z)$), then
  $F \ClosedDay{\bullet} G = \alpha(U F \Day{\bullet} U G)$ and
  $\ClosedDay{j} = \alpha(\Day{j})$ define a residuated commutative
  pomonoid on $\ClosedLowerSet{A}$.
\end{proposition}

\begin{proof}
  We can define an operation
  $\bullet^c : \mathrm{ctxt}(F) \times \mathrm{ctxt}(G) \to
  \mathrm{ctxt}(F \Day{\bullet} G)$ that ``multiplies'' two trees,
  such that
  $\mathit{sum}(c) \bullet \mathit{sum}(d) \leq \mathit{sum}(c
  \bullet^c d)$, using the distributivity. This allows us to show that
  $\alpha$ preserves the monoid operation:
  $\alpha F \ClosedDay{\bullet} \alpha F = \alpha (F \Day{\bullet}
  G)$. With this, we can show that the monotonicity, associativity,
  unit, and commutativity properties of $\Day{\bullet}$ transfer over
  to $\ClosedDay{\bullet}$. The definition of the residual from lower
  sets is already $+$-closed, by distributivity.
\end{proof}

\begin{proposition}\label{prop:closed-monoid-duoidal}
  For a pomonoid $(\lhd, j)$ on $(A, \leq)$, if this satisfies
  $(w \lhd x) + (y \lhd z) \leq (w + y) \lhd (x + z)$ then the Day
  construction
  $F \Day{\lhd} G = \{ z \mid z \leq x \lhd y, x \in F, y \in G \}$ on
  lower sets is $+$-closed when $F$ and $G$ are. We write
  $F \ClosedDay{\lhd} G$ to indicate when we mean this construction as
  an operation on $+$-closed lower sets. If $j + j \leq j$, then the
  Day unit $\Day{j} = \eta(j)$ is also closed and we write it as
  $\ClosedDay{j} \in \ClosedLowerSet{A}$. Together
  $(\ClosedDay{\lhd}, \ClosedDay{j})$ form a pomonoid on
  $(\ClosedLowerSet{A}, \subseteq)$.
\end{proposition}

\begin{proof}
  Since $+$ is entropic over $(\lhd, j)$, the Day monoid $\Day{\lhd}$
  is automatically $+$-closed. The monoid structure directly
  transfers. Similarly, $\eta(j)$ is automatically $+$-closed since
  $j + j \leq j$.
\end{proof}

\begin{remark}
  Generalising the situation for the unit $j$ in Proposition
  \ref{prop:closed-monoid-duoidal}, $\eta(x)$ is closed for any $x$
  such that $x + x \leq x$. Note that if $+$ were a join on
  $(A, \leq)$, then this would automatically be satisfied.
\end{remark}

\begin{remark}
  We have used the same decoration $\ClosedDay{\bullet}$ and
  $\ClosedDay{\lhd}$ for two separate constructions of pomonoids on
  $+$-closed lower sets. We will be careful to distinguish which we
  mean: in our present application, a symmetric operator like
  $\bullet$ will distribute over $+$ and so $\ClosedDay{\bullet}$ will
  be constructed by Proposition \ref{prop:closed-monoid-distrib}; and
  a non-symmetric operator like $\lhd$ will be entropic with respect
  to $+$ and so $\ClosedDay{\lhd}$ will be constructed by Proposition
  \ref{prop:closed-monoid-duoidal}.
\end{remark}

\begin{remark}
  If we have two pomonoids on $(A, \leq)$ that share a unit, then the
  two constructions of units in Propositions
  \ref{prop:closed-monoid-distrib} and
  \ref{prop:closed-monoid-duoidal} will yield the same element of
  $\ClosedLowerSet{A}$.
\end{remark}

\begin{proposition}
  Let $(\bullet, i)$ and $(\lhd, j)$ be pomonoids on $(A, \leq)$ in a
  duoidal relationship, and assume that $(\bullet, i)$ distributes
  over $+$ (as in Proposition \ref{prop:closed-monoid-distrib}) and
  $+$ is entropic over $(\lhd, j)$ (as in Proposition
  \ref{prop:closed-monoid-duoidal}). Then
  $(\ClosedDay{\bullet}, \ClosedDay{i})$ and
  $(\ClosedDay{\lhd}, \ClosedDay{j})$ are in a duoidal relationship on
  $(\ClosedLowerSet{A}, \subseteq)$.
\end{proposition}

\begin{proof}
  The entropic relationship established in Proposition
  \ref{prop:lower-set-duoidal} carries over thanks to the properties
  of $\alpha$ and $U$. The fact that
  $\ClosedDay{i} \subseteq \ClosedDay{j}$ relies on the condition
  $j + j \leq j$ to collapse $+$-contexts of $j$s.
\end{proof}

\subsubsection{Chu Construction}
\label{sec:chu}

To construct suitable MAV-algebras, we use the poset version of the
Chu construction \cite{barr}. The Chu construction builds
$*$-autonomous categories from symmetric monoidal closed categories
with pullbacks. In the partial order case, the requirement for
pullbacks simplifies to binary meets. Therefore, for this section, we
let $(A, \leq, \bullet, i, \rightblackspoon)$ be a partial order with
a residuated pomonoid structure and all binary meets.

\begin{definition}
  Let $k$ be an element of $A$. $\Chu(A, k)$ is the partial order with
  elements pairs $(a^+, a^-)$ such that $a^+ \bullet a^- \leq k$, with
  ordering $(a^+,a^-) \sqsubseteq (b^+, b^-)$ when $a^+ \leq b^+$ and
  $b^- \leq a^-$.
\end{definition}

\begin{proposition}
  $(\Chu(A, k), \sqsubseteq)$ has a $*$-autonomous structure defined
  as:
  \begin{displaymath}
    (a^+, a^-) \otimes (b^+, b^-) = (a^+ \bullet b^+, (b^+ \rightblackspoon a^-) \land (a^+ \rightblackspoon b^-))
    \qquad
    I = (i, k)
    \qquad
    \lnot(a^+,a^-) = (a^-, a^+)
  \end{displaymath}
\end{proposition}

\begin{remark}
  If we choose $k = i$, then $(\Chu(A, i), \sqsubseteq)$ has
  $*$-autonomous structure that satisfies \emph{mix}.
\end{remark}

\begin{proposition}\label{prop:chu-meets}
  If $A$ has binary joins, then $(\Chu(A, k), \sqsubseteq) x$ has
  binary meets, given by
  $(a^+,a^-) \with (b^+,b^-) = (a^+ \land b^+, a^- \lor b^-)$.
\end{proposition}

\begin{remark}
  Since $(\Chu(A, k), \sqsubseteq)$ is a $*$-autonomous partial order,
  then Proposition \ref{prop:chu-meets} also means that $\Chu(A, k)$
  has all binary joins, with
  $(a^+, a^-) \oplus (b^+, b^-) = (a^+ \lor b^+, a^- \land b^-)$.
\end{remark}

We now turn to the self-dual duoidal structure required to interpret
the $\vSeq$ connective. First we transfer pomonoids from $(A, \leq)$
to self-dual pomonoids on $(\Chu(A, k), \sqsubseteq)$ provided they
interact well with $k$:
\begin{proposition}\label{prop:chu-self-dual}
  Let $(\lhd, j)$ be a pomonoid on $(A, \leq)$ such that $(\bullet, i)$ is entropic over 
  $k \lhd k \leq k$ and $j \leq k$, then
  $x ; y = (x^+ \lhd y^+, x^- \lhd y^-)$ and $J = (j, j)$ form a
  self-dual pomonoid on $\Chu(A, k)$.
\end{proposition}

\begin{proof}
  $x;y$ is well-defined because
  $(x^+ \lhd y^+) \bullet (x^- \lhd y^-) \leq (x^+ \bullet x^-) \lhd
  (y^+ \bullet y^-) \leq k \lhd k \leq k$. $J$ is well defined because
  $j \bullet j \leq j \leq k$. The pomonoid laws all transfer directly.
\end{proof}

\begin{remark}
  When $k = j$, the two conditions in the proposition are
  automatically satisfied. Moreover, if $k = i = j$, then not only
  does the $*$-autonomous structure satisfy \emph{mix}, but we also
  have $I =J$.
\end{remark}

Finally, we need to show that if $(\bullet, j)$ is duoidal over
$(\lhd, j)$, then their Chu counterparts are in the same
relationship. Due to the use of residuals in the definition of
$\otimes$, we need the following fact about duoidal residuated
pomonoids:

\begin{lemma}\label{lem:duoidal-residual}
  If $(\bullet, j)$ is duoidal over $(\lhd, j)$ in a partial order
  $(A, \leq)$ and $(\bullet, j)$ has a residual $\rightblackspoon$,
  then
  $(w \rightblackspoon x) \lhd (y \rightblackspoon z) \leq (w \lhd y)
  \rightblackspoon (x \lhd z)$.
\end{lemma}

\begin{remark}
  Lemma \ref{lem:duoidal-residual} is in some sense the
  ``intuitionistic'' version of the duoidal relationship for $\parr$
  arising as the dual of that for $\otimes$ in a $*$-autonomous
  partial order.
\end{remark}

\begin{proposition}
  If $(\bullet, i)$ is entropic over $(\lhd, j)$ on $(A, \leq)$, and
  $(\lhd, j)$ satisfies the conditions of Proposition
  \ref{prop:chu-self-dual}, then $(\otimes, I)$ and $(;, J)$ are in a
  duoidal relationship on $\Chu(A, k)$.
\end{proposition}

\begin{proof}
  For the positive half of the Chu construction, this is a direct
  consequence of the entropy on $(A, \leq)$. For the negative half, we
  use Lemma \ref{lem:duoidal-residual} and the fact that meets are
  always duoidal.
\end{proof}

\subsection{Construction of MAV-algebras}

\begin{theorem}
  If $(F, \leq, \parallel, \lhd, I, +)$ is an MAV-frame, then
  $(\Chu(\ClosedLowerSet{F}, I), \sqsubseteq)$ has the structure of an
  MAV-algebra.
\end{theorem}

\begin{proof}
  Combining all the propositions above, in the special case when some
  of the units collapse.
\end{proof}