### 2024-03-26 doc/sections/mav-cut-elimination.tex
[ agda tex ] more conceptual proof of reflection property

\section{Semantic Cut-Elimination}\label{sec:mav-cut-elimination}

\begin{proposition}
  Let $\Chu(\ClosedLowerSet{\textsc{CFMav}}, \ClosedDay{I})$ be the
  MAV-algebra constructed from the Cut-free proof MAV-frame
  (Proposition \ref{prop:cfmav-frame}), where elements are
  positive/negative pairs of $+$-closed lower sets of formulas.

  \bob{describe the interpretation of atoms}

  The interpretation of SMAV formulas in this MAV-algebra satisfies
  the following two properties for every formula $\vP$:
  \begin{enumerate}
  \item $\eta^+(P) \subseteq \sem{\vP}^-$
  \item $\sem{\vP}^+ \subseteq \eta^+(P) \rightblackspoon \ClosedDay{i}$
%  \item $\forall \vR \in \sem{\vP}^+.~(\vR \vParr \vP) \longrightarrow^* I$
  \end{enumerate}
\end{proposition}

\begin{remark}
  These properties are akin to the \emph{reflect} and \emph{reify}
  functions used in Normalisation by Evaluation (NbE). The
  \emph{reflect} property creates a canonical negative proof of every
  formula assuming that formula (corresponding to $\eta$-expansion in
  NbE. The \emph{reify} property extracts a normal-form proof from the
  semantic model.
\end{remark}

\begin{proof}
  By induction on the structure of the formula $\vP$.
  \begin{itemize}
  \item[$\vUnit$] reflexivity
  \item[A+] reflexivity
  \item[A-] reflexivity
  \item[$\vP \vParr \vQ$] $\eta^+$ preserves monoids (FIXME: proposition)
  \item[$\vP \vTens \vQ$] complicated
  \item[$\vP \vWith \vQ$] $\eta^i$ preserves $+$
  \item[$\vP \vPlus \vQ$] interpreted as a meet, so we reason directly
  \item[$\vP \vSeq \vQ$] $\eta^+$ preserves monoids (FIXME: proposition)
  \end{itemize}
\end{proof}

\subsection{The Okada Property}