### 2024-03-29 doc/sections/mav-cut-elimination.tex
Part of pushing through syntactic stuff

\section{Semantic Cut-Elimination and Proof Normalisation}
\label{sec:mav-cut-elimination}

Let $\Chu(\ClosedLowerSet{\NMAV}, \ClosedDay{I})$ be the MAV-algebra
constructed (\Cref{thm:algebra-from-frame}) from the normal
proof MAV-frame (\Cref{prop:nmav-frame}), where elements
are positive/negative pairs of $+$-closed lower sets of formulas. We
define the valuation of atoms as
$V(\va) = \ChuEmbed(\ClosedLowerEmbed(\vDual\va))$. By
\Cref{thm:soundness}, we have an interpretation of MAV formulas
$\sem{\vP}$ such that if $\vP\vInferFrom*\vUnit$, then
$I \sqsubseteq \sem{\vP}$. We now prove our main proposition about this
interpretation in the MAV-algebra derived from the MAV-frame of normal
proofs \NMAV{} that will allow us to derive the admissibility of all
the non-normal proof rules of MAV.

\begin{proposition}\label[prop]{prop:embedding-sem}
  $\sem{\vP} \sqsubseteq \lnot (\ChuEmbed(\ClosedLowerEmbed(P)))$
\end{proposition}

\begin{proof}
  By \Cref{defn:chu}, this statement is two inclusions
  between pairs of $+$-closed lower sets:
  \begin{enumerate}
    \item $\ClosedLowerEmbed(P) \subseteq \sem{\vP}^-$
    \item $\sem{\vP}^+ \subseteq \ClosedLowerEmbed(P) \rightblackspoon^+ \ClosedDay{I}$
  \end{enumerate}
  We prove the second assuming the first. It suffices to prove that
  $\sem{\vP}^+ \ClosedDay{\bullet} \ClosedLowerEmbed(P) \subseteq
    \ClosedDay{I}$, which follows from the first part and the property
  of all Chu-elements that
  $\sem{\vP}^+ \ClosedDay{\bullet} \sem{\vP}^- \subseteq \ClosedDay{I}$.

  We prove the first part by induction on $\vP$.
  In the cases when $\vP = \vUnit$ or $\vP = \vDual\va$, we already have
  $\sem{\vP}^- = \ClosedLowerEmbed(P)$. When $\vP = \va$, we have
  $\sem{\va}^- = \ClosedLowerEmbed(\vDual\va)
    \rightblackspoon^+ \ClosedDay{I}$. It suffices to prove that
  $\ClosedLowerEmbed(\va) \ClosedDay{\bullet}
    \ClosedLowerEmbed(\vDual\va) \subseteq \ClosedDay{I}$, which
  follows from the preservation of monoid operations by
  $\ClosedLowerEmbed$ and the \cref{rule:AtomAxiom} rule.

  When $\vP = \vQ\vParr\vR$, $\vQ\vWith\vR$, or $\vQ\vSeq\vR$, the result
  follows from preservation of the corresponding monoid structure by
  $\ClosedLowerEmbed$. For example,
  $\ClosedLowerEmbed(\vQ\vParr\vR) \subseteq \ClosedLowerEmbed(\vQ)
    \ClosedDay{\bullet} \ClosedLowerEmbed(\vR) \subseteq \sem{\vQ}^-
    \ClosedDay{\bullet} \sem{\vR}^- = \sem{\vQ\vParr\vR}^-$.

  When $\vP = \vQ\vPlus\vR$, we have
  $\ClosedLowerEmbed(\vQ\vPlus\vR) \subseteq \ClosedLowerEmbed(\vQ)$ and
  $\ClosedLowerEmbed(\vQ\vPlus\vR) \subseteq \ClosedLowerEmbed(\vR)$, by the
  \cref{rule:Left} and \cref{rule:Right} rules. Therefore,
  $\ClosedLowerEmbed(\vQ\vPlus\vR) \subseteq \ClosedLowerEmbed(\vQ) \lor
    \ClosedLowerEmbed(\vR) \subseteq \sem{\vQ}^- \lor \sem{\vR}^- = \sem{\vQ\vPlus\vR}^-$.

  When $\vP = \vQ\vTens\vR$, we have
  $\sem{\vQ\vTens\vR}^- = (\sem{\vQ}^+ \rightblackspoon \sem{\vR}^-) \land
    (\sem{\vR}^+ \rightblackspoon \sem{\vQ}^-)$. We prove inclusion in the
  left-hand side, the right-hand side is similar. The key property we
  need to prove is:
  \begin{equation}\label{eq:interaction-leq}
    \ClosedLowerEmbed(\vQ\vTens\vR) \ClosedDay{\bullet} (\ClosedLowerEmbed(Q) \rightblackspoon^+ \ClosedDay{I})
    \subseteq
    \ClosedLowerEmbed(R)
  \end{equation}
  Using the monoidality and monotonicity of $\va$, this inclusion
  is implied by the following inclusion in $\LowerSet{\NMAV}$:
  \begin{displaymath}
    \LowerEmbed(\vQ\vTens\vR) \Day{\bullet} (U(\ClosedLowerEmbed(Q)) \rightblackspoon \Day{I})
    \subseteq
    \LowerEmbed(R)
  \end{displaymath}
  which follows from the \cref{rule:Switch} rule of MAV and
  calculation. Using \ref{eq:interaction-leq}, and inclusion (ii)
  above, we can prove the inequality we need:
  \begin{displaymath}
    \ClosedLowerEmbed(\vQ\vTens\vR) \ClosedDay{\bullet} \sem{\vQ}^+
    \subseteq
    \ClosedLowerEmbed(\vQ\vTens\vR) \ClosedDay{\bullet} (\ClosedLowerEmbed(Q) \rightblackspoon^+ \ClosedDay{I})
    \subseteq
    \ClosedLowerEmbed(R)
    \subseteq
    \sem{\vR}^-
  \end{displaymath}
  Using the residuation property of $\ClosedDay{\bullet}$ we can conclude.
\end{proof}

\begin{theorem}\label[thm]{thm:cut-elim}
  If $\vP\vInferFrom*\vUnit$ in MAV, then there is a normal proof $\vP\vInferFrom*\vUnit$.
\end{theorem}

\begin{proof}
  By \Cref{thm:soundness}, $\vP\vInferFrom*\vUnit$ in MAV
  implies $I \sqsubseteq \sem{\vP}$. Combined with \Cref{prop:embedding-sem}, we have
  $I \sqsubseteq \lnot \ChuEmbed(\LowerEmbed(\vP)$. By
  \Cref{defn:chu} of the ordering of Chu elements, we have
  $\ClosedLowerEmbed(\vP) \subseteq \ClosedDay{I}$. Since
  $P \in \ClosedLowerEmbed(\vP)$, we have $\vP \in \ClosedDay{I}$, which
  by definition means that there is a normal proof
  $\vP\vInferFrom*\vUnit$.
\end{proof}

Another consequence of \Cref{prop:embedding-sem} is that
MAV is complete for the MAV-frame semantics:

\begin{theorem}\label[thm]{thm:completeness}
  MAV is complete for the MAV-frame semantics: if, for all MAV-frames
  $F$, $\sem{\vUnit} \sqsubseteq \sem{\vP}$ in
  $(\Chu(\ClosedLowerSet{F}, \ClosedDay{I}), \sqsubseteq)$, then
  $\vP\vInferFrom*\vUnit$.
\end{theorem}