### 2025-05-02 journal-version/paper.tex
[ tex ] Ported paper to LMCS class, except for appendix

\documentclass{lmcs}

\keywords{Linear Logic, Deep Inference, Algebraic Semantics, Metatheory}

\usepackage{hyperref}
\usepackage{mathcommand}
\usepackage{graphicx}
\usepackage{cmll}
\usepackage{mathpartir}
\usepackage[all]{xy}
\usepackage[inline]{enumitem}
\usepackage{amssymb}
\usepackage{fdsymbol}
\usepackage{stmaryrd}
\usepackage{parskip}
\usepackage[inline]{enumitem}
\usepackage{xspace}
\usepackage{multirow}
\usepackage{longtable}
\usepackage[UKenglish,all]{foreign}
\usepackage[goodsyntax]{virginialake}
\vllineartrue%
%\usepackage{url}
%\sloppy

\input{local-macros}

\begin{document}

\title{A Semantic Proof of Generalised Cut Elimination for Deep Inference}
\titlecomment{Longer version of the MFPS version}
\thanks{This work was funded by the
  \href{https://www.ukri.org/about-us/epsrc/}{Engineering and Physical
    Sciences Research Council}: Grant number
  \href{https://gow.epsrc.ukri.org/NGBOViewGrant.aspx?GrantRef=EP/T026960/1}{EP/T026960/1,
    \emph{AISEC: AI Secure and Explainable by Construction}.}}

\author[R. Atkey]{Robert Atkey}[a]
\author[W. Kokke]{Wen Kokke}[b]

\address{Mathematically Structured Programming Group\\
  Computer and Information Sciences\\
  University of Strathclyde\\
  Glasgow, Scotland, UK}
\email{robert.atkey@strath.ac.uk}

\address{}
\email{me@wen.works}

% FIXME: shorten this
\begin{abstract}
  \emph{Multiplicative-Additive System Virtual} (MAV) is a logic that
  extends Multiplicative-Additive Linear Logic with a self-dual
  non-commutative operator expressing the concept of ``before'' or
  ``sequencing''. MAV is also an extenson of the the logic Basic
  System Virtual (BV) with additives. Formulas in BV have an appealing
  reading as processes with parallel and sequential composition. MAV
  adds internal and external choice operators. BV and MAV are also
  closely related to Concurrent Kleene Algebras.

  Proof systems for MAV and BV are Deep Inference systems, which allow
  inference rules to be applied anywhere inside a structure. As with
  any proof system, a key question is whether proofs in MAV can be
  reduced to a normal form, removing detours and the introduction of
  structures not present in the original goal. In Sequent Calcluli
  systems, this property is referred to as Cut Elimination. Deep
  Inference systems have an analogous Cut rule and other rules that
  are not present in normalised proofs. Cut Elimination for Deep
  Inference systems has the same metatheoretic benefits as for Sequent
  Calculi systems, including consistency and decidability.

  Proofs of Cut Elimination for BV, MAV, and other Deep Inference
  systems present in the literature have relied on intrincate
  syntactic reasoning and complex termination measures.

  We present a concise semantic proof that all MAV proofs can be
  reduced to a normal form avoiding the Cut rule and other ``non
  analytic'' rules. We also develop soundness and completeness proofs
  of MAV (and BV) with respect to a class of models. We have
  mechanised all our proofs in the Agda proof assistant, which
  provides both assurance of their correctness as well as yielding an
  executable normalisation procedure. Our technique extends to include
  exponentials and the additive units.
\end{abstract}

\maketitle

\input{sections/introduction}
%\input{sections/contribution}
\input{sections/mav-syntax}
\input{sections/mav-semantics}
\input{sections/mav-cut-elimination}
\input{sections/mav-mechanisation}
\input{sections/mav-extensions}

\section*{Acknowledgements}

We would like to thank Ross Horne for helpful comments and pointers to
related work. This work was funded by the
\href{https://www.ukri.org/about-us/epsrc/}{UKRI Engineering and
  Physical Sciences Research Council (EPSRC)}, grant number
\href{https://gow.epsrc.ukri.org/NGBOViewGrant.aspx?GrantRef=EP/T026960/1}{EP/T026960/1}
``\emph{AISEC: AI Secure and Explainable by Construction}''.

\bibliographystyle{alpha}
\bibliography{paper}

% \appendix

% \input{appendix/table-of-statements}

\end{document}