### 2025-05-12 journal-version/paper.tex

\documentclass{lmcs}

\keywords{Linear Logic, Deep Inference, Algebraic Semantics, Metatheory}

\usepackage{hyperref}
\usepackage{mathcommand}
\usepackage{cmll}
\usepackage{mathpartir}
\usepackage[inline]{enumitem}
\usepackage{amssymb}
\usepackage{fdsymbol}
\usepackage{stmaryrd}
\usepackage{xspace}
\usepackage{multirow}
\usepackage{longtable}
\usepackage[UKenglish,all]{foreign}
\usepackage[goodsyntax]{virginialake}
\vllineartrue%
%\usepackage{url}
%\sloppy

\input{local-macros}

\begin{document}

\title{A Semantic Proof of Generalised Cut Elimination for Deep Inference}
\titlecomment{Longer version of the MFPS version}
\thanks{This work was funded by the
  \href{https://www.ukri.org/about-us/epsrc/}{Engineering and Physical
    Sciences Research Council}: Grant number
  \href{https://gow.epsrc.ukri.org/NGBOViewGrant.aspx?GrantRef=EP/T026960/1}{EP/T026960/1,
    \emph{AISEC: AI Secure and Explainable by Construction}.}}

\author[R. Atkey]{Robert Atkey}[a]
\author[W. Kokke]{Wen Kokke}[b]

\address{Mathematically Structured Programming Group\\
  Computer and Information Sciences\\
  University of Strathclyde\\
  Glasgow, Scotland, UK}
\email{robert.atkey@strath.ac.uk}

\address{Well-Typed}
\email{me@wen.works}

% FIXME: shorten this
\begin{abstract}
  We prove cut-elimination properties for some Deep Inference systems
  using semantic techniques.
\end{abstract}

\maketitle

\input{sections/introduction}
\input{sections/mav-syntax}
\input{sections/mav-semantics}
\input{sections/mav-cut-elimination}
\input{sections/mav-mechanisation}
\input{sections/mav-extensions}

\section*{Acknowledgements}

We would like to thank Ross Horne for helpful comments and pointers to
related work. This work was funded by the
\href{https://www.ukri.org/about-us/epsrc/}{UKRI Engineering and
  Physical Sciences Research Council (EPSRC)}, grant number
\href{https://gow.epsrc.ukri.org/NGBOViewGrant.aspx?GrantRef=EP/T026960/1}{EP/T026960/1}
``\emph{AISEC: AI Secure and Explainable by Construction}''.

\bibliographystyle{alpha}
\bibliography{paper}

% \appendix

% \input{appendix/table-of-statements}

\end{document}